---
title: "Review"
format: html
editor: visual
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Instructions

1.  **Clone the Homework Repository**\
    Clone the GitHub repository containing the homework file to your **local machine**.

2.  **Rename the File**\
    Rename the `.qmd` file from `hw01.qmd` to: `hw01_yourusername.qmd`.\

For example: `hw01_semiyari.qmd`

### Working on Your Homework

-   As you complete each part of the homework:

    -   **Stage** and **commit** your `.qmd` file *after each question/sub-question*, such as `1.1`, `1.2`, etc.\
    -   **Render** your `.qmd` file to **HTML only** (no PDF or Word formats).\
    -   **Push** your updates to GitHub after completing each **main question**, such as `1.0`, `2.0`, etc.

> Committing and pushing to GitHub **are graded components** of this assignment.

## Final Checks Before Submission

-   **Re-render** to HTML before your final commit and push.\

-   **Check your output:**\

    -   Make sure **all images display correctly**.\
    -   Avoid including **excessive output or large datasets** unless asked.

## Final Submission

Once you have finished and pushed your final version:

-   Copy the link of your **GitHub Pages site** (where the rendered HTML is published).
-   **Post the link** to the Canvas assignment submission page.

```{r}
# Install pacman if it's not already installed
if (!require("pacman")) install.packages("pacman")

# Use pacman to install (if needed) and load the packages
pacman::p_load(pacman, tidyverse,
               nycflights13)
```

-   Consider the data in the nycflights13 package.

-   **Problem 2**: Is there a relationship between the age of a plane and its delays?

Based on the gg plot, there is no clear visible trend showing that older planes are more delayed. Most of the flights, regardless of age, have arrival delays between 100 and 400 minutes. There are a few outliers closer to 1000 minutes across the ages 10-25 years. Plane ages about 25 years+ do not have these more extreme outliers and still appear to perform similarly to younger planes in terms of delays. 

According to the model, plane age is statistically significant with a p-val < .05, however, the size of the effect is small- for every year on a planes age, the arrival delay decreases by .124 minutes. Additionally, the r-squared value is extremely low, .03%, meaning that only .03% of the variability in arrival delay is explained by plane age. In all, although plane age is statistically significant, its effect is small and the model explains little no none of the variance. Therefore, plane age seems to have no relationship with arrival delays. 

```{r}
library(pacman)
pacman::p_load(pacman, tidyverse,
               nycflights13)

view(flights)
view(planes)
```

```{r}
#merge the data sets and create a variable for the age of the plane
flights_planes <- flights |>
  left_join(planes, by = "tailnum") |>
  mutate(plane_age = year.x - year.y) |>
  filter(!is.na(plane_age))
```

```{r}
ggplot(flights_planes, aes(x = plane_age, y = arr_delay)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE, color = "plum") +
  labs(title = "Plane Age and Arrival Delay",
       x = "Plane Age",
       y = "Delay")

model <- lm(arr_delay ~ plane_age, data = flights_planes)
summary(model)
```

-   **Problem 3**: Find the 10 days of the year that have the highest median departure delay, then select all flights from those 10 days.

10 days of the year with the highest median departure delay: 2013-03-08, 2013-07-01, 2013-07-23, 2013-12-17, 2013-06-13, 2013-05-08, 2013-04-12, 2013-06-25, 2013-06-28, 2013-12-23.

```{r}
library(lubridate)
```

```{r}

#individualize the dates
flights_date <- flights |>
  mutate(date = make_date(year, month, day))

#medians of departure delays and the 10 days of the year with the highest medians 
top_10_days <- flights_date |>
  group_by(date) |>
  summarise(median_dep_delay = median(dep_delay, na.rm = TRUE)) |>
  arrange(desc(median_dep_delay)) |>
  slice_head(n = 10)

#all flights from those 10 days 
flights_top_10_dates <- flights_date |>
  semi_join(top_10_days, by = "date")
```






